---
title: "US gun deaths"
author: "Lars"
date: "20 2 2018"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)

opts_knit$set(root.dir = "~/lars_projects/guns")
opts_chunk$set(fig.path="figs/",
               fig.align='center',
               fig.show='hold',
               results='asis',
               echo = FALSE,
               warning = FALSE,
               message = FALSE)
```

```{r}
library(forcats)
library(stringr)
library(lubridate)
library(tidyverse)
```


## Who commits suicide?

Data from <http://fivethirtyeight.com>
Data from 2012 - 2014?

## Age of victims

```{r, age_density}
full_guns <- read_csv("full_data.csv")


full_guns %>%
  filter(intent %in% c("Suicide", "Homicide", "Accidental")) %>%
  ggplot(aes(age, color=intent)) +
  geom_density() +
  facet_wrap(~sex)
```

What stands out here is the peak in density at late 20s for male homicide victims. A huge increase in risk of homicide after about age seventeen. A similar pattern is seen among female victims, but the peak is much less marked.

Similarly is there a bimodal pattern for suicide victims among males, with a peak at late 20s, and a later peak in 50-55 year olds.


```{r, race_density_intent}
full_guns %>%
  filter(intent %in% c("Suicide", "Homicide", "Accidental")) %>%
  ggplot(aes(age, color=intent)) +
  geom_density() +
  facet_wrap(~race)
```

The striking feature of grouping on race is the early deaths of hispanic and black gun victims. The spike is well younger than 30 years old.

```{r}
tmp <- full_guns %>%
  filter(intent %in% c("Suicide", "Homicide")) %>%
  filter(race %in% c("Black", "White", "Hispanic"))
  
tmp %>%
  group_by(race, intent) %>%
  summarise(count = n()) %>%
  kable()
```


```{r, race_abs_n}
  
  ggplot(tmp, aes(age)) +
  geom_histogram(bins = 40) +
  facet_grid(intent~race)
```

In absolute numbers. The numbers of black himicide victims are really high. So are the numbers of suicide among whites.


